---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# biographR

<!-- badges: start -->
<!-- badges: end -->

The biographR package is designed to help extract structured biographical data from unstructured text using ChatGPT's API. 

## Installation

You can install the latest version of biographR using the following code:

``` r
remotes::install_github("SamuelFrederick/biographR")
```

## Example

Say we have a biography for an individual named John Smith. The biography might look something like: 

```
John Smith graduated from Nowhere College with a B.A. in 1962. 
He then went on to receive a Ph.D. from Nonexistent University. 
At the same time, his wife, Sally Smith, was earning her M.D. from Invisible University.
```

It would be challenging to extract structured biographical data about John Smith from this text without reading the text and getting the information by hand. Instead, we can use the `get_bio()` function from the `biographR` package, which calls ChatGPT's API to extract the data for us. 


```{r example}
library(biographR)
get_bio(bio = "John Smith graduated from Nowhere College with a B.A. in 1962. He then went on to receive a Ph.D. from Nonexistent University. At the same time, his wife, Sally Smith, was earning her M.D. from Invisible University.", 
        bio_name = "John Smith", 
        prompt_fields = c("college", "graduate_school", 
                          "highest_level_of_education", "gender",
                          "married"))
```

We can also customize the output further using the `prompt_fields_formats` and `prompt_fields_values` arguments. Additionally, while the package defaults to GPT 3.5, GPT 4 often appears to perform somewhat better in extracting subtler information (e.g., gender from pronouns).

```{r example 2}
get_bio(bio = "John Smith graduated from Nowhere College with a B.A. in 1962. He then went on to receive a Ph.D. from Nonexistent University. At the same time, his wife, Sally Smith, was earning her M.D. from Invisible University.", 
        bio_name = "John Smith", 
        prompt_fields = c("college", "graduate_school", 
                          "highest_level_of_education", "gender",
                          "married"), 
        prompt_fields_formats = list(college = "{SCHOOL} - {DEGREE}", 
                                    graduate_school = "{SCHOOL} - {DEGREE}", 
                                    highest_level_of_education = "{DEGREE}"), 
        prompt_fields_values = list(married = c("Yes", "No")), 
        openai_model = "gpt-4")
```





